---
title: "Inclinación del Arbolado Público en Mendoza"
output: html_notebook
---
#Libraries
```{r}
# For manipulating the datasets
library(dplyr)
library(readr)
# For plotting correlation matrix
library(ggcorrplot)
library(ggplot2)
# Machine Learning library
library(rpart)
library(caret)
```
# GET THE DATA
## Load the dataset
```{r}
trainset <- readr::read_csv("./arbolado-mza-dataset.csv",
                              col_types = cols(
  id = col_integer(),
  especie = col_character(),
  ultima_modificacion = col_character(),
  altura = col_character(),
  circ_tronco_cm = col_double(),
  diametro_tronco = col_character(),
  long = col_double(),
  lat = col_double(),
  seccion = col_integer(),
  nombre_seccion = col_character(),
  area_seccion = col_double(),
  inclinacion_peligrosa = col_integer()
))

trainset<-trainset %>% mutate(inclinacion_peligrosa=ifelse(inclinacion_peligrosa=='1','si','no'))
```
## Analize the dataset
#Distribución de la inclinación peligrosa
```{r}
trainset %>%  group_by(inclinacion_peligrosa) %>% summarise(total=n())
```
#Inclinación por sección
```{r}
trainset_seccion_inclinacion<-trainset %>%  group_by(seccion,inclinacion_peligrosa) %>% summarise(total=n())
trainset_seccion_inclinacion

#Visualización
ggplot(trainset)+
  geom_bar(aes(x=as.factor(inclinacion_peligrosa),fill=as.factor(inclinacion_peligrosa)   ))+
  theme_bw()+
  facet_wrap(~seccion)
```
#Inclinación por especie
```{r}
trainset_especie_inclinacion<-trainset %>%  group_by(especie,inclinacion_peligrosa) %>% summarise(total=n())
trainset_especie_inclinacion

#Visualización
ggplot(trainset)+
  geom_bar(aes(x=as.factor(inclinacion_peligrosa),fill=as.factor(inclinacion_peligrosa)   ))+
  theme_bw()+
  facet_wrap(~especie)# Aca estamos indicando que queremos que diferencie por especie
```
# CLEAN, PREPARE & MANIPULATE THE DATA
```{r}
##Inclinación peligrosa as factor
trainset$inclinacion_peligrosa <-as.factor(trainset$inclinacion_peligrosa)
##Eliminate some variables
trainset <- trainset %>% select(-ultima_modificacion,-nombre_seccion,-area_seccion)
```
```{r}
pp <- trainset %>% group_by(especie) %>% summarise(total = n()) %>% filter(total > 1)
trainset <- trainset %>% filter(especie %in% pp$especie)
```
```{r}
names(trainset) %>% as.data.frame()
```
# TRAIN THE MODEL
## Split train and test
```{r}
trainIndex <- createDataPartition(as.factor(trainset$inclinacion_peligrosa), p=0.70, list=FALSE)
data_train <- trainset[ trainIndex,]
data_test <-  trainset[-trainIndex,]
colnames(data_train) <- make.names(colnames(data_train))
colnames(data_test) <- make.names(colnames(data_test))
```
### Plot class distribution in train
```{r}
data_train  %>% group_by(inclinacion_peligrosa) %>% summarise(total=n()) %>%
  ggplot()+
  geom_col(aes(x=inclinacion_peligrosa,y=total,fill=inclinacion_peligrosa))+
  theme_classic()
```
### Plot class distribution in test
```{r}
data_test  %>% group_by(inclinacion_peligrosa) %>% summarise(total=n()) %>%
  ggplot()+
  geom_col(aes(x=inclinacion_peligrosa,y=total,fill=inclinacion_peligrosa))+
  theme_classic()
```

## Train model
```{r}
ctrl_fast <- trainControl(method="cv", 
                     repeats=1,
                     number=5, 
                   #  summaryFunction=twoClassSummary,
                     verboseIter=T,
                     classProbs=F, #savePredictions = T,
                     allowParallel = TRUE)
```

```{r}

ctrl_fast$sampling<-"up"

svmGrid <-  expand.grid(sigma= c(0.001,0.0001,0.00001), 
                        C = c(1,2,4,8,16,32,64,80,100,120) 
                        )

#svmGrid <-  expand.grid(C= c(100), sigma = c(1))


train_formula<-formula(inclinacion_peligrosa~.)
rfFitupsam<- train(train_formula,
               data = data_train,
               #method = "rf",   # Radial kernel
               #method = "xgbTree",
               method = "rpart",
               #tuneLength = 9,
               #tuneGrid = svmGrid,
               #preProcess=c("scale","center"),
               #metric="ROC",
               #weights = model_weights,
               trControl = ctrl_fast)

#plot(rfFitupsam)
rfFitupsam
#rfFitupsam$finalModel
```
# TEST THE DATA
```{r}
predsrfprobsamp=predict(rfFitupsam,data_test)
# use for regresion
cbind(predsrfprobsamp,data_test$inclinacion_peligrosa)
#confusionMatrix(as.factor(predsrfprobsamp %>% round()),as.factor(data_test$inclinacion_peligrosa))

confusionMatrix(predsrfprobsamp,as.factor(data_test$inclinacion_peligrosa))
```
#Load the dataset again
```{r}
trainset1 <- readr::read_csv("./arbolado-mza-dataset.csv",
                              col_types = cols(
  id = col_integer(),
  especie = col_character(),
  ultima_modificacion = col_character(),
  altura = col_character(),
  circ_tronco_cm = col_double(),
  diametro_tronco = col_character(),
  long = col_double(),
  lat = col_double(),
  seccion = col_integer(),
  nombre_seccion = col_character(),
  area_seccion = col_double(),
  inclinacion_peligrosa = col_integer()
))

trainset1<-trainset1 %>% mutate(inclinacion_peligrosa=ifelse(inclinacion_peligrosa=='1','si','no'))
trainset1 <- trainset1 %>% select(-diametro_tronco,-ultima_modificacion,-nombre_seccion,-area_seccion)
trainset1$inclinacion_peligrosa <- as.factor(trainset1$inclinacion_peligrosa)

```
#Train the model again with all the data
```{r}
ctrl_fast <- trainControl(method="cv", 
                     repeats=1,
                     number=5, 
                   #  summaryFunction=twoClassSummary,
                     verboseIter=T,
                     classProbs=F,
                     allowParallel = TRUE)

ctrl_fast$sampling<-"up"

svmGrid <-  expand.grid(sigma= c(0.001,0.0001,0.00001), 
                        C = c(1,2,4,8,16,32,64,80,100,120) 
                        )

train_formula<-formula(inclinacion_peligrosa~.)
rfFitupsam<- train(train_formula,
               data = trainset1,
               #method = "rf",   # Radial kernel
               #method = "xgbTree",
               method = "rpart",
               #tuneLength = 9,
               #tuneGrid = svmGrid,
               #preProcess=c("scale","center"),
               #metric="ROC",
               #weights = model_weights,
               trControl = ctrl_fast)

#plot(rfFitupsam)
rfFitupsam
#rfFitupsam$finalModel

```
#Load the data-test
```{r}
data_pred <-  readr::read_csv("./arbolado-mza-dataset-test.csv",col_types = cols(
  id = col_integer(),
  especie = col_character(),
  ultima_modificacion = col_character(),
  altura = col_character(),
  circ_tronco_cm = col_double(),
  diametro_tronco = col_character(),
  long = col_double(),
  lat = col_double(),
  seccion = col_integer(),
  nombre_seccion = col_character(),
  area_seccion = col_double()
))

data_pred <- data_pred %>% select(-ultima_modificacion,-nombre_seccion,-area_seccion)
```
#Prdict the results
```{r}
pred=predict(rfFitupsam,data_pred)
```
#Save results
```{r}
pred_val = ifelse(pred=="si",1,0)
submission<-data.frame(id=data_pred$id,inclinacion_peligrosa=pred_val)
readr::write_csv(submission,"./arbolado-mza-dataset-envio-nuevo.csv")
head(submission)
```

